Dynamic Load Balancing for\\Heterogeneous Systems

### 2011-12-05 paper/paper.tex
Fleshed out architecture section.

\subsection{CPU Architecture}
The CPU architecture is designed as a low latency architecture.  To facilitate
the needs for low latency application performance, a series of hierarchical 
caches have been designed to hide memory access costs, the most common cause
of latency in single-threaded applications on modern architectures.  Each core
of the CPU typically only runs one or two hardware-enabled threads to reduce
cache pollution, thus avoiding latency.  An example CPU architecture is given
by Figure~\ref{fig:cpu_arch}.

The use of the Single Instruction Multiple Data (SIMD) paradigm enables increased
performance on these devices by increasing data throughput for data parallel
applications.

\subsection{GPU Architecture}
The GPU almost acts a foil to the CPU architecture.  Instead of optimizing the
architecture of the device to increase performance of a single thread, the GPU
architecture is specifically designed for large throughput performance.

The memory hierarchy for the GPU architecture is given by Figure~\ref{fig:gpu_arch}.

>>>>

\subsection{CPU Architecture}
The CPU architecture is designed as a low latency architecture.  To facilitate
the needs for low latency application performance, a series of hierarchical 
caches have been designed to hide memory access costs, the most common cause
of latency in single-threaded applications on modern architectures.  Each core
of the CPU typically only runs one or two hardware-enabled threads to reduce
cache pollution, thus avoiding latency.  An simplistic example CPU architecture is given
by Figure~\ref{fig:cpu_arch}.

The use of the Single Instruction Multiple Data (SIMD) paradigm enables increased
performance on these devices by increasing data throughput for data parallel
applications.  However, optimal performance of these types of instructions often
involves complex compilation techniques and hand-tuning of the assembly.

\subsection{GPU Architecture}
The GPU almost acts a foil to the CPU architecture.  Instead of optimizing the
architecture of the device to increase performance of a single thread, the GPU
architecture is specifically designed for large throughput performance.  Large
numbers of threads can be easily created an run on a GPU.  Large register files
enable tens of thousands of threads to be run on today's GPU architecture with
very little penalty for context switching.

Groups of proccessing units, Streaming Multiprocessors (SMs), share a single instruction
dispatch unit and operate in lock step.  The threads themselves run on each SM
in a group, called a warp, typically of 32 threads.  Because of these architectual
features divergent branching in a warp (caused by if-else clauses) greatly degrade application
performance.

The memory architecture for the GPU is very different than that found on the CPU. 
Instead of a large hierarchy of caches to hide memory latency for single thread
performance the latecy to memory is instead hidden through aggressive context
switching.  Newer GPU architectures may contain primative caching schemes.
Each SM on the GPU also contains a local memory memory space which can accessed 
at much lower latency .  This memory space is typically used as a user-managed
cache. The memory hierarchy for the GPU architecture is given by Figure~\ref{fig:gpu_arch}.
